---
title: "Bertrand Tornil's Resume"
author: Bertrand Tornil
date: "`r Sys.Date()`"
output:
  pagedown::html_resume:
    # set it to true for a self-contained HTML page but it'll take longer to render
    self_contained: false
    css: ['css/custom_resume.css', 'css/styles_html.css', 'resume']
# uncomment this line to produce HTML and PDF in RStudio:
#knit: pagedown::chrome_print
params:
  doctype: "HTML"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  results='asis',
  echo = FALSE
)
```

# Aside

![Dr Mustache](https://www.tornil.net/cv/images/mustache.svg){width=80%}

## Contact Info {#contact}

-   <i class="fa fa-envelope"></i> [bertrand.tornil\@gmail.com](mailto:bertrand.tornil@gmail.com){.email}
-   <i class="fa fa-globe-americas"></i> [www.tornil.net](https://www.tornil.net/)
-   <i class="fa fa-linkedin"></i> [bertrandtornil](https://www.linkedin.com/in/bertrandtornil/)
-   <i class="fa fa-twitter"></i> [twitter.com/bobuss](https://twitter.com/bobuss)
-   <i class="fa fa-github"></i> [github.com/bobuss](https://github.com/bobuss)
-   For more information, please contact me via email.

```{r download}
# if doctype is html add link to download a pdf version
if(params$doctype == "HTML"){
  cat("[<i class='fas fa-download'></i> Download as a PDF](https://www.tornil.net/cv/Bertrand_Tornil_Resume.pdf)")
}
```

## Skills {#skills}

-   WebService Architectures
-   DevOps
-   Open Source

## Disclaimer {#disclaimer}

- Made w/ [**pagedown**](https://github.com/rstudio/pagedown).
- Source code: [**github repo**](https://github.com/bobuss/cv).

Last updated on `r Sys.Date()`.
<br>
<br>


# Main

## Bertrand Tornil {#title}

### Software Engineer, PhD

I deliver available, scalable and observable web services.

My days are full of Ansible, MariaDB, Redis, Python, Orchestrator, Kafka, Azure, Prometheus, grafana,
OpenTelemtry, Solr, JVM, shell, RedHat.
I also use Terraform, Kubernetes, docker, Clickhouse, Consul, and many other cool tools...
Sometimes, I am a frontend hobbyist, currently playing with vue.js and Audioworklet API
I used to be on-call, and I liked it!


## Education {data-icon="graduation-cap" data-concise="true"}

### Université Paul Sabatier

PhD in Information Systems (Doctorat d'informatique)

Toulouse, France

2006

Thesis (french): [Interaction gestuelle et haptique, pour plus d'utilisabilité et d'accessibilité](https://www.tornil.net/these)

### Université Paul Sabatier

MA in Computer Information (Master Image & Language)

Toulouse, France

2001

### Université de Limoges (Licence et Maitrise d'informatique)

BA in Computer Science

Limoges, France

1999

## Professional Experience {data-icon="suitcase"}

### Principal Software Engineer

[Itron](https://www.itron.com/)

Remote, USA

Present - 2016

-   Technical swiss knife : SmartCities solution architecture
-   [StreetLight.Vision](https://www.itron.com/na/solutions/what-we-enable/smart-cities/central-management-software)
-   Duties and achievements:
    -   Architecture to allow Highly Availability on Premise and Cloud environments
    -   Choice of new tools and open source software: git, Prometheus, Grafana, Kafka, Redis, orchestrator
    -   MariaDB High-Availability (10sec failover) based on Orchestrator, HAProxy, Consul
    -   Ansible automation
    -   Transition to Azure DevOps Services: IAAS and Pipelines for builds and deployements
    -   Kubernetes, Terraform
    -   Design and deploy the Monitoring and Alerting stack
    -   DevOps evangelism: Continuous Integration, Continuous Deployment
    -   Technological watch, everyday

### Data Scientist, Web architect, Lead Dev, Release Manager

[IsCool Entertainment - Paris](https://www.iscoolentertainment.com)

Paris, France

2016 - 2009

-   2015 - 2016 : Data Scientist
    -   Postgresql, w/ PipelineDB, CitusDB
    -   InfluxDB
    -   Google BigQuery
    -   Kubernetes
-   2013 - 2015 : Lead Dev on [Lord of the Dumbs](https://www.lordofthedumbs.com/fr/)
    -   Python (Pyramid, SQLAlchemy, uWSGI, supervisord, Fabric)
    -   postgresql, redis
    -   RabbitMQ
    -   RESTful webservices
    -   AngularJS
-   2012 : R&D Transversal Projects:
    -   RealTime dashboards (node.js / mongodb / html5)
    -   QA and Continuous Delivery Processes and Tools (Jenkins, Tests, Capistrano, Monitoring)
-   2011 : Lead Dev & Release Manager on ColorParty: (up to 150,000 Monthly Active Users)
    -   Shoots engine: 3Mshoots / day
    -   Players Logbook Consolidation in best-effort
    -   Business metrics, Users profiles, Messaging on Redis
-   2009 - 2010 : Lead Dev & Release Manager on IsCool: 1,500,000 Monthly Active Users, 80M Pages / day
    -   Tokens sharing engine: still in production, 2G daily shared points
    -   Redis deployment
    -   Symfony2 Transition (see [Iscool Story. Symfony Live 2011 - Paris](https://fr.slideshare.net/bertrandtornil/symfony-live-2011))
    -   Release process
    -   SVN to GIT Transition
    -   Makefiles (assets compilation, i18n)
    -   Capistrano, EC2, S3
-   Development architecture on EC2
-   Agile: scrum (certified), kanban and XP notions
-   DevOps: 1 release per day, Monitoring via Pinba and Zabbix
-   Management: 10-20 persons team, hiring, trainees

### Core Developper

[Dailymotion - Paris](https://www.dailymotion.com)

Paris, France

2009 - 2008

-   PHP / Mysql / memcache / git
-   Localization
-   Revenue sharing (ads servers)
-   audio & video fingerprinting and videos publication process

### Web consultant

[ALTEN - Boulogne Billancourt](https://www.alten.fr)

Paris, France

2008 - 2007

-   PHP, Mysql, Javascript
-   Web Developpment: [Virgin Mobile](https://www.virginmobile.fr)
-   R&D: [Cellfish media](https://www.cellfishmedia.fr) (Lagardère Active)
-   Web Developpment: [caradisiac.com](https://www.caradisiac.com)

### R&D engineer

UIXPERTS - Paris

Remote, France

2006

-   PHP, Mysql, Javascript

### Web Developper

[WS Interactive - Toulouse](https://www.ws-interactive.fr)

Toulouse France

2006

-   PHP, Mysql, Javascript


## Research Experience {data-icon="laptop"}

### Graduate Research Assistant

[IRIT & Université Paul Sabatier - Toulouse](https://www.irit.fr)

Toulouse, France

2009 - 2000

-   Search domains: HCI, Tactile and force feedback Interfaces, AI
-   Teaching: CAML, C, Pascal, Delphi, Algorithms
-   [Publications](https://tornil.net/bibliography)

## Teaching Experience {data-icon="chalkboard-teacher"}

### AWS, scalability and DevOps (french)

[INSSET, Master Cloud Computing](http://apache.insset.u-picardie.fr/formations/master-informatique/master-ccm)

Saint-Quentin, France

2016-2013

-   [AWS, scalability and DevOps (french)](https://www.tornil.net/INSSET_AWS/index.html)

### Cloud et architectures scalables (french)

[EPITA](https://www.epita.fr)

Paris, France

2011

-   [Cloud et architectures scalables (french)](https://www.tornil.net/cours/epita2011.html)

## Selected Publications {data-icon="file"}

### Iscool Story

Symfony Live

Paris, France

2011

Alexandre Rodière, **Bertrand Tornil**

<i class="fa fa-slideshare"></i> [slideshare.net/bertrandtornil/symfony-live-2011](https://fr.slideshare.net/bertrandtornil/symfony-live-2011).

### Music haptic: Musical harmony notions for all with a force feedback mouse and a spatial representation

International Computer Music Conference

Barcelona, Spain

2005

**Bertrand Tornil**, Nadine Baptiste-Jessel

### Use of force feedback pointing devices for blind users

8th ERCIM Workshop User Interfaces For All

Vienna, Austria

2004

**Bertrand Tornil**, Nadine Baptiste-Jessel
